---
title: "CMDA-2006 Homework"
subtitle: "Homework"
author: "Your name here"
date: "Due"
output:
  pdf_document:
    highlight: haddock
keep_tex: no
number_sections: no
html_document:
  df_print: paged
geometry: margin = 0.5in
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
editor_options:
  chunk_output_type: console
documentclass: article
urlcolor: blue
---
  
<!-- The above is set to automatically compile to a .pdf file.   -->
<!-- It will only succeed if LaTeX is installed. -->
  
<!-- If you absolutely can't get LaTeX installed and/or working, then you can compile to a .html first,  -->
<!-- by clicking on the arrow button next to knit and selecting Knit to HTML. -->

<!-- You must then print you .html file to a .pdf by using first opening it in a web browser and then printing to a .pdf -->


```{r setup, include=FALSE}
# This is the setup chunk
#  Here you can set global options for the entire document

library(knitr) # I recommend doing this

# Although you can call functions from a library using the following notation
#  without loading the entire library.
knitr::opts_chunk$set(echo = TRUE, 
                      comment = NA, # Recommended
                      fig.path = "./figures/",  # Store all figures here in relative path (make the folder first)
                      fig.align = "center",
                      fig.width = 7,
                      fig.height = 7,
                      message = FALSE, # Turn off load messages
                      warning = FALSE # Turn off warnings
                      )

```

\clearpage

```{r include=FALSE}
# You should not echo this chunk.
# include=FALSE does more than echo=FALSE, it actually does: echo=FALSE, results='hide', message=FALSE, warning=FALSE, fig.show='hide'

# You should set your working directory at the very beginning of your R Markdown file
# setwd("~/Dropbox/teaching/SU19/CMDA_3654/homework/homework1/")

# In linux ~/ is shorthand for /home/username/
# You should type things out properly for your system
# Mac: /Users/username/Documents/CMDA3654/Lectures/Lecture_03/.../
# Windows: C:/Users/username/Documents/etc/Lecture/Lecture_03/.../


```

<!-- ---------------------------------------------------------------------------------------------------- -->
<!-- ---------------- Problems start below these lines ----------------------------------------- -->
<!-- ---------------------------------------------------------------------------------------------------- -->
```{r}
setwd("C:/Users/jlutz/OneDrive/Documents/School/Spring Sem 2022/CMDA-STAT/Homework/HW_7")
```

# Problem 1

Perform a one-way analysis of variance to test whether there is a significant difference in the average
leaf sizes under the four growing conditions. Use alpha = 0.05.
## Part a

Example Chunk Below

```{r}
Leaf_A <- c(27.7619, 27.8523, 21.3495, 31.9616, 19.4623, 12.2804, 21.0508, 19.5074, 26.2808, 26.1466)

Leaf_B <- c(4.2460, 14.1577, 7.0279, 7.0698, 0.8091, 13.9385, 11.0130, 10.9680, 6.9112, 9.6041)

Leaf_C <- c(15.5070, 5.0473, 18.3020, 16.0436, 10.2601, 19.0571, 17.1826, 16.6510, 18.8472, 12.4234)

Leaf_D <- c(33.0101, 44.9680, 34.2074, 28.9766, 42.9229, 36.6827, 32.7229, 34.5668, 28.7695, 36.6952)

Leaf_Data <- c(Leaf_A, Leaf_B, Leaf_C, Leaf_D)
leaves <- c(rep("Leaf A", 10), rep("Leaf B", 10), rep("Leaf C", 10), rep("Leaf D", 10))
anovafit0 <- aov(Leaf_Data ~ leaves)
print(anova0 <- anova(anovafit0))
```

## Part b

What conclusions can you reach concerning the effect of growing conditions on the average leaf size?

Since our p-value, 1.736e-13 < alpha = 0.05, we reject the null hypothesis. We have enough statistical evidence to conclude that there is a significant difference concerning the effect of growing conditions on the average leaf size. 


## Part c

Perform a one-way analysis of variance to test whether there is a significant difference in the average
nicotine contents under the four growing conditions. Use alpha = 0.05.

```{r}
Nico_A <- c(10.0655, 9.4712, 9.1246, 11.3652, 11.3976, 11.2936, 10.6805, 8.1280, 10.5066, 10.6579)

Nico_B <- c(8.5977, 8.1299, 11.3401, 9.3470, 9.3049, 10.0193, 9.5843, 6.4603, 8.2589, 5.0106)

Nico_C <- c(6.7865, 10.9249, 11.3878, 9.7022, 8.0371, 10.7187, 11.2352, 7.7079, 7.5653, 9.0922)

Nico_D <- c(9.9553, 5.8495, 10.3005, 9.7140, 10.7543, 8.0262, 13.1326, 11.8559, 11.3345, 10.4763)

Nico_Data <- c(Nico_A, Nico_B, Nico_C, Nico_D)
titles <- c(rep("A Nico", 10), rep("B Nico", 10), rep("C Nico", 10), rep("D Nico", 10))

anovafit1 <- aov(Nico_Data ~ titles)
print(anova1 <- anova(anovafit1))
```

## Part d

What conclusions can you reach concerning the effect of growing conditions on the average nicotine
content?

Since our p-value, 0.117 > alpha = 0.05, we fail to reject the null hypothesis. We do not have enough statistical evidence to conclude that there is a significant difference concerning the effect of growing conditions on the average nicotine content. 

## Part e

Based on the conclusions you reached in parts (b) and (d), do you think the testimony of the tobacco
companies scientists is supported by this experiment? Justify your conclusions.

I believe the testimony of the scientists is not supported by this experiment. From the one-way analysis of variance on the average leaf sizes from the four growing conditions, we determined that their is statistical evidence to conclude that their is a significant difference concerning the effect of growing conditions on the average leaf size. In the the one-way analysis on the growing conditions on the average nicotine content, we determined there is not sufficient statistical evidence to conclude significant difference concerning the growing conditions on the average nicotine content. 
Since there was evidence for significant difference concerning the effect of growing conditions on the average leaf size, but not on the average nicotine content, we have evidence to support that the companies scientists testimony was not supported by this experiment. 

## Part f

If you observe significant F -tests in either parts (b) or (d). Carry out the pair-wise tests to determine
which mean(s) are different.

```{r}
pairwise.t.test(Leaf_Data, leaves, conf.level = 0.95)
```


## Part g

Do the nicotine content data suggest violations of the AOV conditions? If you determine that the
conditions are not met, perform an alternative analysis, and compare your results to the earlier analysis.
Additionally if you observe a significant test, do the pair-wise comparisons using a non-parametric
procedure with a Bonferroni correction.

```{r}
#install.packages("car")
library(car)
bartlett.test(Nico_Data ~ titles)
leveneTest(Nico_Data ~ titles)
fligner.test(Nico_Data ~ titles)
```

Since all three tests agree that we do not reject the null hypothesis, we will assume that all the variances are equal, we do not have suggestions of violations of the AOV conditions.

```{r}
pairwise.t.test(Nico_Data, titles, p.adjust.method = "bonf")
```


# Problem 2
```{r}
selenium <- read.csv("selenium.csv", header = T)
```

## Part a

First, conduct the analysis of variance on this data set.
```{r}
selenium$Meat <- factor(selenium$Meat)
anovafit2 <- aov(Selenium ~ Meat, data = selenium)
print(anova2 <- anova(anovafit2))
```

## Part b
Specifically identify the null and alternative hypothesis statements for the contrasts that we are testing
in the above questions.

For Question 1 :
    H0: VEN <= SQU 
    HA: VEN > SQU
For Question 2:
    H0: RRB == NRB
    HA: RRB != NRB
For Question 3: 
    H0: (VEN + SQU) == (RRB + NRB)
    HA: (VEN + SQU) != (RRB + NRB)

## Part c
Using the Bonferroni method, find the contrast estimates, along with their 95% simultaneous confidence intervals and the adjusted p-values for the above research questions. This can be done using a
single contrast matrix in R. Display the results, and provide plots of the confidence intervals.

```{r}
library(DescTools)
pairwise.t.test(selenium$Selenium, selenium$Meat, conf.level = 0.95, p.adjust.method = "bonf")
t <- 4
k <- t * (t - 1)/2
df.within <- anova2[2,"Df"]
library(multcomp)
comparisons <- glht(anovafit2, linfct = mcp(Meat = "Tukey"))
plot(confint(comparisons, calpha=qt(1-0.05/2/k, df.within)),
     main = "Bonferroni 95% family-wise confidence level")
```

## Part d

Don’t forget to answer the original research questions using part (c).

Since our p-value 7.684e-14 < alpha = 0.05, we reject h). We have sufficient statistical evidence to conclude that there is a significant difference.

## Part e

Finally, use Tukeys method to conduct a post-hoc analysis of all of the pairwise contrast estimates,
their 95% confidence intervals. Display the results, and provide a plot of the confidence intervals.
Comment on any interesting results that you see.

```{r}
PostHocTest(anovafit2, method = "bonferroni")
plot(confint(comparisons, calpha=qt(1-0.05/2/k, df.within)),
     main = "Bonferroni 95% family-wise confidence level")
```



